---
title: "Wine Review Analysis"
output: html_notebook
---

##### Data Environment Set-Up
```{r} 
library(Amelia)
library(tidyverse)
library(tm)
library(wordcloud)
library(readr)
library(tidytext)
library(SnowballC)
library(wordcloud)
library(ggmap)
library(maps)
library(mapdata)
library(tree)
library(randomForest)
library(gmodels)
library(caret)
library(pROC)
library(cvTools)
library(C50)
library(mltools)


wine.og <- read.csv("https://wine-review-dataset.s3.amazonaws.com/wine_dataset.csv")
wine <- wine.og

```
  

## Data Cleaning
```{r}
summary(wine)
```
  
##### Replace "" with NAs
```{r}
# Points
ggplot(data = wine) + geom_histogram(aes(points), binwidth = 1)
# Price
ggplot(data = filter(wine, price < 500)) + geom_histogram(aes(price))
expensive.wine <-  filter(wine, price >= 500)
  # only 104 pricey bottles - remove from wine dataset
wine <- filter(wine, price < 500)
# Country
wine[which(wine$country == ""), "country"] <- NA
# Designation
wine[which(wine$designation == ""), "designation"] <- NA
# Province
wine[which(wine$province == ""), "province"] <- NA
# Region 1
wine[which(wine$region_1 == ""), "region_1"] <- NA
# Region 2
wine[which(wine$region_2 == ""), "region_2"] <- NA
# Taster Name
wine[which(wine$taster_name == ""), "taster_name"] <- NA
# Twitter Handle - REMOVE
colnames(wine)
wine <- wine[,-11]
```
  
##### Remove Duplicates 
```{r}
wine <- wine[!(duplicated(wine[,c("title", "winery", "variety", "designation", "country", "taster_name")])),]
  # Removed 754 duplicated 
```
  
##### Examine Missing Values
```{r}
missmap(wine)
  # Region 2 - mostly NA, don't use in analysis
  # Designation - half NA, can't impute
  # Taster Name - half NA, can't impute
  # Region 1 - some NAs
```
  
  
## Text Mining
  
##### Description Corpus
```{r}
# Enter each description as a work in Corpus
wine.desc <- Corpus(VectorSource(wine$description))
# Make all words lowercase, remove punctuation, remove stop words
wine.desc.clean <- wine.desc %>%
  tm_map(content_transformer(tolower)) %>%
  tm_map(removeWords, stopwords('english')) %>%
  tm_map(removeWords, 'wine') %>%
  tm_map(removePunctuation) %>%
  tm_map(removeNumbers)
wine.desc.clean[100]$content
wine[100, "description"]
```
  
##### Most Frequent Words
```{r}
dtm <- DocumentTermMatrix(wine.desc.clean)
DF <- tidy(dtm)
names(DF) <- c('document','word','freq')
DF <- arrange(DF, desc(freq))
head(DF)
# Bar Plot of dtm
DF %>% filter(freq>3) %>% ggplot(aes(word,freq)) + geom_col() + xlab(NULL)+ coord_flip()
# Word Cloud
wordcloud(wine.desc.clean,
          scale=c(5,0.5),
          max.words=100, 
          min.freq=4,
          random.order=FALSE,
          rot.per=.35,
          use.r.layout=FALSE,
          colors=brewer.pal(8,"Dark2"))
```
  
##### Tastes & Aromas
```{r}
# Tastes Files
fruit.tastes <- read.table("fruit.txt")
fruit <- as.character(fruit.tastes$V1)
fruit <- tolower(fruit)
# Add flavors to wine dataset
wine$flavors <- NA
for(i in 1:nrow(wine)) {
  desc <- wine.desc.clean[i]$content
  desc.words <- strsplit(desc, " ")
  desc.words <- unlist(desc.words)
  for(j in 1:length(desc.words)) {
    if(!is.na(match(desc.words[j],fruit))) {
      wine[i, "flavors"] <- desc.words[j]
      break
    }
  }
}
wine$flavors <- as.factor(wine$flavors)

# Aromas File
aromas <- read.table("aromas.txt")
aroma <- as.character(aromas$V1)
aroma <- tolower(aroma)
# Add aromas to wine dataset
wine$aromas <- NA
for(i in 1:nrow(wine)) {
  desc <- wine.desc.clean[i]$content
  desc.words <- strsplit(desc, " ")
  desc.words <- unlist(desc.words)
  for(j in 1:length(desc.words)) {
    if(!is.na(match(desc.words[j],aroma))) {
      wine[i, "aromas"] <- desc.words[j]
      break
    }
  }
}
wine$aromas <- as.factor(wine$aromas)
```
  
  
## Graphical Analysis
  
##### Wine Variety -> Color
```{r}
# With variety
wine.variety <- wine %>%
  group_by(variety) %>%
  summarize(count = n())
wine.variety <- arrange(wine.variety, desc(count))
##wine.variety[41:50,]
wine$color <- NA
wine <- mutate(wine, color = ifelse(variety == "Pinot Noir", "red", 
                                    ifelse(variety == "Chardonnay", "white", 
                                           ifelse(variety == "Cabernet Sauvignon", "red",
                                                  ifelse(variety == "Red Blend", "red", 
                                                         ifelse(variety == "Riesling", "white", 
                                                                ifelse(variety == "Sauvignon Blanc", "white", 
                                                                       ifelse(variety == "Ros????", "white", 
                                                                              ifelse(variety == "Zinfandel", "red", 
                                                                                     ifelse(variety == "Malbec", "red", 
                                                                                            ifelse(variety == "Sangiovese", "red",
                                                                                                   ifelse(variety == "Nebbiolo", "red",
                                                                                                          ifelse(variety == "Portuguese Red", "red", 
                                    ifelse(variety == "White Blend", "white", 
                                           ifelse(variety == "Sparkling Blend", "white",
                                                  ifelse(variety == "Tempranillo", "red",
                                                         ifelse(variety == "Rh????ne-style Red Blend", "red", 
                                                                ifelse(variety == "Pinot Gris", "white", 
                                                                       ifelse(variety == "Cabernet Franc", "red",
                                                                              ifelse(variety == "Gr????ner Veltliner", "white",
                                                                                     ifelse(variety == "Champagne Blend", "white",
                                                                                            ifelse(variety == "Pinot Grigio", "white", 
                                    ifelse(variety == "Viognier", "white", 
                                           ifelse(variety == "Portuguese White", "white", 
                                                  ifelse(variety == "Gew????rztraminer", "white",
                                                         ifelse(variety == "Shiraz", "red", 
                                                                ifelse(variety == "Gamay", "red",
                                                                       ifelse(variety == "Petite Sirah", "red", 
ifelse(variety == "Bordeaux-style White Blend", "white",
       ifelse(variety == "Grenache", "red", 
              ifelse(variety == "Tempranillo Blend", "red",
                    ifelse(variety == "Barbera", "red",
                            ifelse(variety == "Sangiovese Grosso", "red",
                                   ifelse(variety == "Glera", "white",
                                          ifelse(variety == "Carmen????re", "red", 
                                                 ifelse(variety == "Chenin Blanc", "white",
                                                        ifelse(variety == "Corvina, Rondinella, Molinara", "red",
                                                               ifelse(variety == "Port", "red",
ifelse(variety == "Albari????o", "white",
       ifelse(variety == "Rh????ne-style White Blend", "white",
              ifelse(variety == "Pinot Blanc", "white",
                     ifelse(variety == "Garnacha", "red",
                            ifelse(variety == "Moscato", "white",
                                   ifelse(variety == "Nero d'Avola", "red",
                                          ifelse(variety == "Sauvignon", "white",
                                                 ifelse(variety == "Verdejo", "white",
                                                        ifelse(variety == "Petit Verdot", "red",
                                                               ifelse(variety == "Aglianco", "red",  NA))))))))))))))))))))))))))))))))))))))))))))))))

wine$color <- as.factor(wine$color)
ggplot(data=wine %>% group_by(color) %>% summarize(count = n()),
       aes(x=color, y=count, fill=color)) +
  geom_col()
```
  
```{r}
# Scatter Plot
ggplot(data = filter(wine, !is.na(color))) + 
  geom_point(mapping = aes(x = price, y = points, color = color)) +
  geom_smooth(mapping = aes(x = price, y = points)) + 
  facet_wrap(~color)
```
  
```{r}
# Histogram
priceRW <- ggplot(data = filter(wine, !is.na(color))) + 
  geom_histogram(mapping = aes(price, fill = color, color = "grey"), position = "identity", alpha = 1/2) + 
  ylab("frequency") + ggtitle("Comparison of Price for White and Red Wine")
priceRW + scale_fill_manual(values = c("red", "white")) + theme(legend.title = element_blank())
pointsRW <- ggplot(data = filter(wine, !is.na(color))) + 
  geom_histogram(mapping = aes(points, fill = color, color = "grey"), position = "identity", alpha = 1/2, binwidth = 1) + 
  ylab("frequency") + ggtitle("Comparison of Points for White and Red Wine")
pointsRW + scale_fill_manual(values = c("red", "white")) + theme(legend.title = element_blank())
```
  
##### Taste & Aroma
```{r}
# Flavor
wine.flavors <- wine %>%
  filter(!is.na(flavors)) %>%
  filter(!is.na(color)) %>%
  group_by(flavors) %>%
  summarize(count = n(),
            white.count = sum(ifelse(color == "white", 1, 0)),
            red.count = sum(ifelse(color == "red", 1, 0)))
ggplot(data = filter(wine.flavors, count > 1500)) + 
  geom_col(aes(x = flavors, y = count,
               fill = ifelse(red.count > white.count, "red", "white"))) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  ylab("frequency") + labs(fill="Most Common Color") + ggtitle("Most Common Fruit Flavors")
```
  
```{r}
# Aroma
wine.aromas <- wine %>%
  filter(!is.na(aromas)) %>%
  group_by(aromas) %>%
  summarize(count = n())
ggplot(data = filter(wine.aromas, count > 1500)) + 
  geom_col(aes(x = aromas, y = count)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  ylab("frequency") + labs(fill="Most Common Color") + ggtitle("Most Common Aromas")
```
  
##### Taster Name
```{r}
wine.taster <- wine %>%
  filter(!is.na(taster_name)) %>%
  group_by(taster_name) %>%
  summarize(frequency = n(),
            avg_points = mean(points))
ggplot(data = wine.taster) + 
  geom_col(aes(x = taster_name, y = frequency, fill = avg_points)) + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  ggtitle("Frequency and Average Points for Each Taster")
```
  
  
## Classification Models
  
##### Build Complete Datasets
```{r}
colnames(wine)
wine.complete <- wine[,-c(1, 3, 4, 8, 9, 11, 13)]
  # Remove X, description, designation, region 1 and 2, title
colnames(wine.complete)
missmap(wine.complete)

# Flavor and aroma
wine.flavaro <- wine.complete[, c(2:3, 6:9)]
nlevels(wine.flavaro$aromas)
summary(wine.flavaro$aromas)
arrange(wine.aromas, desc(count))[31:40,]
aromas.keep <- c("oak", "sweet", "pepper", "vanilla", "chocolate", "spicy", "earthy", "mineral", "herb", "wood",
                 "leather", "floral", "smoky", "toast", "honey", "tobacco", "rose", "coffee",
                 "violet", "almond", "mint", "caramel", "sugar", "tea", "meaty", "grassy", "bacon", "buttery",
                 "burnt", "pine", "rubber")
arrange(wine.flavors, desc(count))[21:30,]
flavors.keep <- c("cherry", "apple", "blackberry", "plum", "rasberry", "lemon", "peach", "pear", "strawberry", "orange",
                 "currant", "pineapple", "grapefruit", "lime", "apricot", "cranberry", "melon", "blueberry", "grape",
                 "tangerine", "olive", "coconut", "tomato", "prune", "raisin", "pomegranate", "nectarine", "boysenberry", "mango")
wine.flavaro <- subset(wine.flavaro, !is.na(match(aromas, aromas.keep)))
wine.flavaro <- subset(wine.flavaro, !is.na(match(flavors, flavors.keep)))
wine.flavaro <- wine.flavaro[complete.cases(wine.flavaro),-3]
wine.flavaro$aromas <- as.character(wine.flavaro$aromas)
wine.flavaro$aromas <- as.factor(wine.flavaro$aromas)
wine.flavaro$flavors <- as.character(wine.flavaro$flavors)
wine.flavaro$flavors <- as.factor(wine.flavaro$flavors)

# Location
wine.locations <- wine %>%
  filter(!is.na(country)) %>%
  group_by(country) %>%
  summarize(count = n())
wine.locations <- arrange(wine.locations, desc(count))
countries.keep <- c("US", "France", "Italy", "Spain", "Portugal", "Chile", "Argentina", "Austria", "Australia", "Germany")
wine.loc <- wine[,c(2,5:6)]
wine.loc <- subset(wine.loc, !is.na(match(country, countries.keep)))
wine.loc$country <- as.character(wine.loc$country)
wine.loc$country <- as.factor(wine.loc$country)
nlevels(wine.loc$country)
summary(wine.loc$country)
```
  
#### Color Prediction Random Forest
  
Prep the DataFrame
```{r}
# Make dataframe with no NAs
wine.col <- wine[,c(5:6, 16)]
wine.col <- wine.col[complete.cases(wine.col),]
# Split into test and train sets
set.seed(12345)
wine.col.rand <- wine.col[order(runif(nrow(wine.col))), ]
wine.col.train <- wine.col.rand[1:round(nrow(wine.col.rand)*(2/3)), ]
wine.col.test  <- drop_na(wine.col.rand[round(nrow(wine.col)*(2/3))+1:nrow(wine.col.rand), ])
```
  
Visualize One Decision Tree
```{r}
# Build simple decision tree
color.tree <- C5.0(wine.col.train[,-3], wine.col.train$color)
plot(color.tree)
```
  
Random Forest Model
```{r}
color.rf <- randomForest(wine.col.train[,-3], wine.col.train$color, ntree = 500, mtry = 2)

# Variable Importance
varImpPlot(color.rf,type=2, main="Variable Importance in Predicting Color")

# Model Performance
p.color <- predict(color.rf, wine.col.test[,-3])
CrossTable(wine.col.test$color, p.color,
           prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE,
           dnn = c('actual color', 'predicted color'))
```
  
Random Forest Model with Cross Validation
```{r}
# With CV
k <- 8
folds <- cvFolds(NROW(wine.col.rand), K=k)
wine.col.rand$holdoutpred <- rep(0,nrow(wine.col.rand))
for(i in 1:k){
  wine.col.train.cv <- wine.col.rand[folds$subsets[folds$which != i], ]
  wine.col.test.cv <- wine.col.rand[folds$subsets[folds$which == i], ]
  
  color.tree.cv <- C5.0(wine.col.train.cv[,-3], wine.col.train.cv$color)
  p.color.cv <- predict(color.tree.cv, wine.col.test.cv[,-3])
  
  wine.col.rand[folds$subsets[folds$which == i], ]$holdoutpred <- p.color.cv
}
levels(wine.col.rand$holdoutpred) <- c("red", "white")
table(wine.col.rand$holdoutpred,wine.col.rand$color)

# Model Performance
CrossTable(wine.col.rand$color, wine.col.rand$holdoutpred,
           prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE,
           dnn = c('actual color', 'predicted color'))
```
  
### Flavor & Aroma Linear Model
  
One-Hot Encoding
```{r}
contrasts(wine.flavaro$flavors)
contrasts(wine.flavaro$aromas)
contrasts(wine.flavaro$color)
contrasts(wine.flavaro$color)<-matrix(c(1, 0),nrow=2)
colnames(contrasts(wine.flavaro$color)) <-"red"
```
  
Predicting Price
```{r}
lm.flavaro.price <- lm(price ~ ., data = wine.flavaro)
summary(lm.flavaro.price)
lm.flavaro.price.step <- step(lm.flavaro.price, trace = F)
summary(lm.flavaro.price.step)
```
  
Predicting Points
```{r}
lm.flavaro.points <- lm(points ~ ., data = wine.flavaro)
summary(lm.flavaro.points)
lm.flavaro.points.step <- step(lm.flavaro.points, trace = F)
summary(lm.flavaro.points.step)
```
  
Linear Regression Model with Cross Validation
```{r}
# With CV
set.seed(12345)
wine.flavaro.rand <- wine.flavaro[order(runif(25597)), ]
k <- 10
folds <- cvFolds(NROW(wine.flavaro.rand), K=k)
pmse.price.result<-rep(0,k)
pmse.points.result<-rep(0,k)
for (i in c(1:k)){
  # generate training sets and test sets from original data:
  wine.fa.train.cv <- wine.flavaro.rand[folds$subsets[folds$which != i], ]
  wine.fa.test.cv <- wine.flavaro.rand[folds$subsets[folds$which == i], ]
  
  # Build model with train set:
  lm.fa.price.cv <- lm(price ~ ., data = wine.fa.train.cv)
  lm.fa.points.cv <- lm(points ~ ., data = wine.fa.train.cv)
  
  # First, how to predict with lm models:
  price.pred<-predict(lm.fa.price.cv,newdata=wine.fa.test.cv) 
  points.pred<-predict(lm.fa.points.cv,newdata=wine.fa.test.cv) 
  
  # Next, compute PMSE:
  pmse.price<-mse(price.pred, wine.fa.test.cv$price)
  pmse.points<-mse(points.pred, wine.fa.test.cv$points)
  
  # Add the PMSE for this run into your vector to stor PMSE
  pmse.price.result[i]<-pmse.price
  pmse.points.result[i]<-pmse.points
}

plot(pmse.price.result,type='b',col='blue',xlab="Index", ylab="PMSE")
abline(mean(pmse.price.result), 0, col='black',xlab="Index", ylab="PMSE")
title(main="PMSE of Price Model")

plot(pmse.points.result,type='b',col='blue',xlab="Index", ylab="PMSE")
abline(mean(pmse.points.result), 0, col='black',xlab="Index", ylab="PMSE")
title(main="PMSE of Points Model")
```
  
#### Location Linear Model
  
Country One-Hot Encoding
```{r}
contrasts(wine.loc$country)
contrasts(wine.loc$country)<-matrix(c(1,0,0,0,0,0,0,0,0,0, 
                                      0,1,0,0,0,0,0,0,0,0,
                                      0,0,1,0,0,0,0,0,0,0, 
                                      0,0,0,1,0,0,0,0,0,0,  
                                      0,0,0,0,1,0,0,0,0,0,
                                      0,0,0,0,0,1,0,0,0,0,
                                      0,0,0,0,0,0,1,0,0,0,
                                      0,0,0,0,0,0,0,1,0,0,
                                      0,0,0,0,0,0,0,0,1,0,
                                      0,0,0,0,0,0,0,0,0,0),nrow=10) # changing value of columns
colnames(contrasts(wine.loc$country)) <-matrix(c("Argentina","Australia","Austria","Chile","France", "Germany", "Italy", "Portugal", "Spain"),ncol=9) # changing titles of columns 
```
  
Predicting Price
```{r}
lm.loc.price <- lm(price ~ ., data = wine.loc)
summary(lm.loc.price)
```
  
Predicting Points
```{r}
lm.loc.points <- lm(points ~ ., data = wine.loc)
summary(lm.loc.points)
```
  
Linear Regression Model with Cross Validation
```{r}
# With CV
set.seed(12345)
wine.loc.rand <- wine.loc[order(runif(106097)), ]
k <- 20
folds <- cvFolds(NROW(wine.loc.rand), K=k)
pmse.price.result<-rep(0,(k))
pmse.points.result<-rep(0,(k))
for (i in c(1:k)){
  # generate training sets and test sets from original data:
  wine.loc.train.cv <- wine.loc.rand[folds$subsets[folds$which != i], ]
  wine.loc.test.cv <- wine.loc.rand[folds$subsets[folds$which == i], ]
  
  # Build model with train set:
  lm.loc.price.cv <- lm(price ~ ., data = wine.loc.train.cv)
  lm.loc.points.cv <- lm(points ~ ., data = wine.loc.train.cv)
  
  # First, how to predict with lm models:
  price.pred<-predict(lm.loc.price.cv,newdata=wine.loc.test.cv) 
  points.pred<-predict(lm.loc.points.cv,newdata=wine.loc.test.cv) 
  
  # Next, compute PMSE:
  pmse.price<-mse(price.pred, wine.loc.test.cv$price)
  pmse.points<-mse(points.pred, wine.loc.test.cv$points)
  
  # Add the PMSE for this run into your vector to stor PMSE
  pmse.price.result[i]<-pmse.price
  pmse.points.result[i]<-pmse.points
}

plot(pmse.price.result,type='b',col='blue',xlab="Index", ylab="PMSE")
abline(mean(pmse.price.result), 0, col='black',xlab="Index", ylab="PMSE")
title(main="PMSE of Price Model")

plot(pmse.points.result,type='b',col='blue',xlab="Index", ylab="PMSE")
abline(mean(pmse.points.result), 0, col='black',xlab="Index", ylab="PMSE")
title(main="PMSE of Points Model")
```


